Memories of migrations past: Sociality and cognition in dynamic, seasonal environments

### 2021-07-13 text/SocialMemoryMigration.tex
ms modifications and figure improvements

@@

\emph{to do.}

@@

Within a given year, the resource is entirely symmetric: $h_y(x,t) = h_y(-x, \tau-t)$. However, in scenarios exploring climate change we allow the peaks to vary with drift and stochasticity according to: $m_x(y) \sim {N}(\mu_x + \beta_x\,y, \sigma_x)$ and $m_t(y) \sim {N}(\mu_t + \beta_t\,y, \sigma_t)$, where the $\mu$, $\beta$ and $\sigma$ terms are the mean, slope and variance, respectively, for the location and time duration of the pulse. Thus, if $\beta=0$ and $\sigma=0$, the conditions are constant across years, if $\beta_x > 0$ there is a drift of the resource towards extremes, if $\beta_t < 0$ there is a shift towards earlier resource pulses. These trends mirror the pole-ward shift of peak resources and the earlier spring phenology occurring with a warming global climate. The spatial scales ($s_x$ and $s_t$) remain constant in all of our simulations.

@@

	\centering
	\begin{tabular}{ll|l}
		\hline\hline
		\multicolumn{3}{l}{\textbf{Memory migration model}}\\ 
		& $\epsilon$ & Diffusion \\ 
		& $\alpha$ & Forage following \\ 
		& $\beta$ & Strength of sociality \\ 
		& $\lambda$ & Spatial scale of sociality \\ 
		& $\kappa$ & Proportion of reference versus working memory \\ 
		& $x_1$, $x_2$ & location of population centroids in summer and winter \\ 
		& $t_1$, $dt_1$ & start and duration of summer season \\ 
		& $t_2$, $dt_2$ & start and duration of winter season \\ 
		& $\kappa$ & proportion memory allocated to reference \\
		&& (long-term) memory vs. working (short-term) memory \\ 
		\multicolumn{3}{l}{\textbf{Resource dynamics}} \\ 
		& $\tau$ & duration of period (year) \\ 
		& $m_x$, $-m_x$ & spatial coordinate of resource peak for summer and winter \\ 
		& $m_t$, $\tau - m_t$ & timing of resource peak for the summer and winter \\ 
		& $\sigma_x$, $\sigma_t$ & time duration and and spatial scale of resource pulse \\ 
		& $dx$, $dt$ & rate of change of peak location and timing of resource \\ 
		& $\psi_x$, $\psi_t$ & standard deviation of peak location and timing \\ 
		\multicolumn{3}{l}{\textbf{Metrics}}\\ 
		& $MM_x$ & spatial migration mismatch \\ 
		& $MM_t$ & temporal migration mismatch \\ 
		& $TE$ & total mismatch \\ 
		& $FE$ & foraging efficiency \\ 
		& $SA$ & spatial adaptation index \\ 
		\hline\hline
	\end{tabular}
\end{table}

@@

\begin{figure}
\includegraphics[width=\textwidth]{figures/Stochasticity.png}

\caption{\label{fig_stochasticity} Foraging efficiency across various values of reference memory $\kappa$ (x-axis) for increasing amounts of interannual stochasticity ($\psi_x$, panels left to right), and two values of spatial scale $\lambda = 80$ and $120$. For the stochastic processes ($\psi_x > 0$), the process was run 90 times, and the mean and standard error of the foraging efficiency over a 50 year post-stabilization period is presented (points and bars). In these scenarios, the resource following parameter $\alpha = 100$, the social attraction $\beta = 400$ and diffusion $\epsilon = 4$.}

@@

EG, WFF, GCC and RSC, were supported in part by NSF award DMS 1853478. EG and WFF were further supported by NSF grant XXXXXXXXXX. The authors are grateful to Quentin Read for computational support and advice for cluster-based multi-node multi-core model runs.

>>>>

%\linenumbers

@@

@@

Within a given year, the resource is entirely symmetric: $h_y(x,t) = h_y(-x, \tau-t)$. However, in scenarios exploring climate change we allow the peaks to vary with drift and stochasticity according to: $m_x(y) \sim {\cal N}(\mu_x + \gamma_x\,y, \sigma_x)$ and $m_t(y) \sim {N}(\mu_t + \gamma_t\,y, \sigma_t)$, where the $\mu$, $\gamma$ and $\sigma$ terms are the mean, slope and variance, respectively, for the location and time duration of the pulse. Thus, if $\gamma=0$ and $\sigma=0$, the conditions are constant across years, if $\beta_x > 0$ there is a drift of the resource towards extremes, if $\beta_t < 0$ there is a shift towards earlier resource pulses. These trends model the pole-ward shift of peak resources and the earlier spring phenology occurring with a warming global climate. The spatial scales ($s_x$ and $s_t$) remain constant in all of our simulations.

@@

	\centering
	\begin{tabular}{ll|l}
		\hline\hline
		\multicolumn{3}{l}{\textbf{Memory migration model}}\\ 
		& $\epsilon$ & Diffusion \\ 
		& $\alpha$ & Forage following \\ 
		& $\beta$ & Strength of sociality \\ 
		& $\lambda$ & Spatial scale of sociality \\ 
		& $\kappa$ & Proportion of reference versus working memory \\ 
		& $x_1$, $x_2$ & location of population centroids in summer and winter \\ 
		& $t_1$, $dt_1$ & start and duration of summer season \\ 
		& $t_2$, $dt_2$ & start and duration of winter season \\ 
		& $\kappa$ & proportion memory allocated to reference \\
		&& (long-term) memory vs. working (short-term) memory \\ 
		\multicolumn{3}{l}{\textbf{Resource dynamics}} \\ 
		& $\tau$ & duration of period (year) \\ 
		& $m_x$, $-m_x$ & spatial coordinate of resource peak for summer and winter \\ 
		& $m_t$, $\tau - m_t$ & timing of resource peak for the summer and winter \\ 
		& $\sigma_x$, $\sigma_t$ & time duration and and spatial scale of resource pulse \\ 
		& $\gamma_x$, $\gamma_t$ & rate of change of peak location and timing of resource \\ 
		& $\psi_x$, $\psi_t$ & standard deviation of peak location and timing \\ 
		\multicolumn{3}{l}{\textbf{Metrics}}\\ 
		& $MM_x$ & spatial migration mismatch \\ 
		& $MM_t$ & temporal migration mismatch \\ 
		& $TE$ & total mismatch \\ 
		& $FE$ & foraging efficiency \\ 
		& $SA$ & spatial adaptation index \\ 
		\hline\hline
	\end{tabular}
\end{table}

@@

\begin{figure}
\includegraphics[width=\textwidth]{figures/stochasticity.png}

\caption{\label{fig_stochasticity} Foraging efficiency (FE) across various values of reference memory $\kappa$ (x-axis) for increasing amounts of interannual stochasticity ($\psi_x$, panels left to right), and two values of sociality spatial scale $\lambda = 80$ and $120$. For the stochastic processes ($\psi_x > 0$), the process was run 90 times for values of $\kappa$. Points represent the means of the FE, and shaded areas represent a loess smooth of FE against $\kappa$. In these scenarios, the resource following parameter $\alpha = 100$, the social attraction $\beta = 400$ and diffusion $\epsilon = 4$.}

@@

EG, WFF, GCC and RSC, were supported in part by NSF award DMS 1853478. EG and WFF were further partially supported by NSF grant IIBR 1915347. The authors are grateful to Quentin Read for computational support and advice for cluster-based multi-node multi-core model runs.
